### 2020-04-25 ms-a4-2_non-linear-relationships.rmd
Added additional resources pages

---
title: "MS-4.2 Types of Relationships"
output:
  html_document:
    code_folding: "hide"
---

# [Syllabus](https://syllabus.nesa.nsw.edu.au/assets/mathematics_standard/mathematics-standard-stage-6-syllabus-2017.pdf) {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

Students:

* use an exponential model to solve problems
  + graph and recognise an exponential function in the form $y=a^x$ and $y=a^{-x} (a>0)$ using technology
  + interpret the meaning of the intercepts of an exponential graph in a variety of contexts
  + construct and analyse an exponential model to solve a practical growth or decay problem
* construct and analyse a quadratic model to solve practical problems involving quadratic functions of expressions of the form $y=ax^2+bx+c$, for example, braking distance against speed
  + recognise the shape of a parabola and that it always has a turning point and an axis of symmetry
  + graph a quadratic function using technology
  + interpret the turning point and intercepts of a parabola in a practical context
  + consider the range of values for $x$ and $y$ for which the quadratic model makes sense in ap ractical context
* recognise that reciprocal functions of the form $y=\frac{k}{x}$, where $k$ is a constant, represent inverse variation, identify the rectangular hyperbolic shape of these graphs and their important features
  + use a reciprocal model to solve practical inverse variation problems algebraically and graphically, eg the amount of pizza received when sharing a pizza between increasing numbers of people
:::

## Hide

#  {.tabset .tabset-fade .tabset-pills}

## Exponentials {.tabset .tabset-fade .tabset-pills}

### Functions

```{r}
library("ggplot2")
eq = function(x){2^x}
ggplot(data.frame(x=c(-3, 3)), aes(x=x)) + 
  stat_function(fun=eq)
```

### Modelling

## Quadratics {.tabset .tabset-fade .tabset-pills}

### Functions

### Modelling

## Reciprocals {.tabset .tabset-fade .tabset-pills}

### Functions

### Modelling

## Inverse variation

## Miscellaneous problems