### 2021-01-20 ms-a1_formulae-and-equations.rmd
Changed icons and theme.

---
title: "MS-A1 Formulae and Equations"
output:
  html_document:
    toc: true
    toc_float: true
---

# Syllabus [![](open.svg){width=16}](https://syllabus.nesa.nsw.edu.au/assets/mathematics_standard/mathematics-standard-stage-6-syllabus-2017.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Hide

## Show

::: {.syllabus}

Students:

* review substitution of numerical values into linear and non-linear algebraic expressions and equations
  + review evaluating the subject of a formula, given the value of other pronumerals in the formula
  + change the subject of a formula
  + solve problems involving formulae, including calculating distance, speed and time (with change of units of measurement as required) or calculating stopping distances of vehicles using a suitable formula
* develop and solve linear equations, including those derived from substituting values into a formula, or those developed from a word description
* calculate and interpret blood alcohol content (BAC) based on drink consumption and body weight
  + use formulae, both in word form and algebraic form, to calculate an estimate for blood alcohol content $(BAC)$, including $BAC_{Male}=\frac{10N-7.5H}{6.8M}$ and $BAC_{Female}=\frac{10N-7.5H}{5.5M}$ where $N$ is the number of standard drinks consumed, $H$ is the number of hours of drinking, and $M$ is the person's weight in kilograms.
  + determine the number of hours required for a person to stop consuming alcohol in order to reach zero BAC, eg using the formula $time=\frac{BAC}{0.015}$
  + describe the limitations of methods estimating BAC
* calculate required medication dosages for children and adults from packets, given age or weight, using Fried’s, Young’s or Clark’s formula as appropriate
  + Fried's formula: Dosage for children 1-2 years $=\frac{\text{age (in months)}\ \times\ \text{adult dosage}}{150}$
  + Young's formula: Dosage for children 1-12 years $=\frac{\text{age of child (in years)}\ \times\ \text{adult dosage}}{\text{age od child (in years)}\ +\ 150}$
  + Clark's formula: Dosage $=\frac{\text{weight in kg}\ \times\ \text{adult dosage}}{70}$

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

# Substitution [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/a_substitution.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Hide

## Show

::: {.syllabus}

**Syllabus**

* review substitution of numerical values into linear and non-linear algebraic expressions and equaitons

:::

### Why algebra?

**Algebra** is used Mathematicians to communicate mathematical ideas in a convenient way. **Variables** (letters and symbols) are used to represent unknown quantities whose value can change depending on the situation.

### Expressions

We can think of expressions as number crunching machines. Numbers are put in and then a related number is produced.

For example the expression $a^2+1$ starts with an input $a$, squares it and adds 1.

If the number -5 is fed into the machine, the machine **substitutes** the number 5 in place of a and then **evaluates** the result:

$$\begin{aligned}&a^2+1\\=\ &(\color{red}{-5})^2+1\\=\ &25+1\\ =\ &26\end{aligned}$$

::: {.important}

**Important Note**

To avoid making mistakes it is good practice to **always** substitute into brackets.

:::

### Examples {.tabset .tabset-fade .tabset-pills}

#### Example 1 {.tabset .tabset-fade .tabset-pills}

##### Question

::: {.example}

Evaluate the following, given $x=4$

a) $\quad2x-3$
b) $\quad\sqrt{6x+12}$

:::

##### Show Solution (a)

Evaluate the following, given $x=4$

a) $\quad2x-3$
b) $\quad\sqrt{6x+12}$

::: {.solution}

$\begin{aligned}&2x-3\\=\ &2(\color{red}{4})-3\\=\ &8-3\\=\ &5\end{aligned}$

:::

##### Show Solution (b)

Evaluate the following, given $x=4$

a) $\quad2x-3$
b) $\quad\sqrt{6x+12}$

::: {.solution}

$\begin{aligned}&\sqrt{6x+12}\\=\ &\sqrt{6(\color{red}{4})+12}\\=\ &\sqrt{24+12}\\=\ &\sqrt{36}\\=\ &6\end{aligned}$

:::

#### Example 2 {.tabset .tabset-fade .tabset-pills}

##### Question

::: {.example}

Evaluate $\quad a^2-5b+c\quad$ given $a=3$, $b=-3$, $c=7$

:::

##### Show Solution

Evaluate $\quad a^2-5b+c\quad$ given $a=3$, $b=-3$, $c=7$

::: {.solution}

$\begin{aligned}&a^2-5b+c\\=\ &(\color{red}{3})^2-5(\color{red}{-3})+(\color{red}{7})\\=\ &9+15+7\\=\ &31\end{aligned}$

:::