### 2021-01-21 ms-a1_formulae-and-equations.rmd
Added MS-A1_c and MS-A1_d

---
title: "MS-A1 Formulae and Equations"
output:
  html_document:
    toc: true
    toc_float: true
---

# Syllabus [![](open.svg){width=16}](https://syllabus.nesa.nsw.edu.au/assets/mathematics_standard/mathematics-standard-stage-6-syllabus-2017.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

Students:

* review substitution of numerical values into linear and non-linear algebraic expressions and equations
  + review evaluating the subject of a formula, given the value of other pronumerals in the formula
  + change the subject of a formula
  + solve problems involving formulae, including calculating distance, speed and time (with change of units of measurement as required) or calculating stopping distances of vehicles using a suitable formula
* develop and solve linear equations, including those derived from substituting values into a formula, or those developed from a word description
* calculate and interpret blood alcohol content (BAC) based on drink consumption and body weight
  + use formulae, both in word form and algebraic form, to calculate an estimate for blood alcohol content $(BAC)$, including $BAC_{Male}=\frac{10N-7.5H}{6.8M}$ and $BAC_{Female}=\frac{10N-7.5H}{5.5M}$ where $N$ is the number of standard drinks consumed, $H$ is the number of hours of drinking, and $M$ is the person's weight in kilograms.
  + determine the number of hours required for a person to stop consuming alcohol in order to reach zero BAC, eg using the formula $time=\frac{BAC}{0.015}$
  + describe the limitations of methods estimating BAC
* calculate required medication dosages for children and adults from packets, given age or weight, using Fried’s, Young’s or Clark’s formula as appropriate
  + Fried's formula: Dosage for children 1-2 years $=\frac{\text{age (in months)}\ \times\ \text{adult dosage}}{150}$
  + Young's formula: Dosage for children 1-12 years $=\frac{\text{age of child (in years)}\ \times\ \text{adult dosage}}{\text{age od child (in years)}\ +\ 150}$
  + Clark's formula: Dosage $=\frac{\text{weight in kg}\ \times\ \text{adult dosage}}{70}$

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

## Hide

# Substitution [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/a_substitution.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* review substitution of numerical values into linear and non-linear algebraic expressions and equaitons

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Why algebra?

**Algebra** is used Mathematicians to communicate mathematical ideas in a convenient way. **Variables** (letters and symbols) are used to represent unknown quantities whose value can change depending on the situation.

### Expressions

We can think of expressions as number crunching machines. Numbers are put in and then a related number is produced.

For example the expression $a^2+1$ starts with an input $a$, squares it and adds 1.

If the number -5 is fed into the machine, the machine **substitutes** the number 5 in place of a and then **evaluates** the result:

$$\begin{aligned}&a^2+1\\=\ &(\color{red}{-5})^2+1\\=\ &25+1\\ =\ &26\end{aligned}$$

::: {.important}

**Important Note**

To avoid making mistakes it is good practice to **always** substitute into brackets.

:::

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Evaluate the following, given $x=4$

a) $\quad2x-3$
b) $\quad\sqrt{6x+12}$

:::

###### Solution (a)

Evaluate the following, given $x=4$

a) $\quad2x-3$
b) $\quad\sqrt{6x+12}$

::: {.solution}

$\begin{aligned}&2x-3\\=\ &2(\color{red}{4})-3\\=\ &8-3\\=\ &5\end{aligned}$

:::

###### Solution (b)

Evaluate the following, given $x=4$

a) $\quad2x-3$
b) $\quad\sqrt{6x+12}$

::: {.solution}

$\begin{aligned}&\sqrt{6x+12}\\=\ &\sqrt{6(\color{red}{4})+12}\\=\ &\sqrt{24+12}\\=\ &\sqrt{36}\\=\ &6\end{aligned}$

:::

##### Example 2 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Evaluate $\quad a^2-5b+c\quad$ given $a=3$, $b=-3$, $c=7$

:::

###### Solution

Evaluate $\quad a^2-5b+c\quad$ given $a=3$, $b=-3$, $c=7$

::: {.solution}

$\begin{aligned}&a^2-5b+c\\=\ &(\color{red}{3})^2-5(\color{red}{-3})+(\color{red}{7})\\=\ &9+15+7\\=\ &31\end{aligned}$

:::

##### Hide

## Practice {.tabset .tabset-fade .tabset-pills}

### Questions

If $a=4$, $b=-1$, and $c=-3$, evaluate:

a) $\quad3a$
b) $\quad2a-3b$
c) $\quad2ac$
d) $\quad ac-3c$
e) $\quad\dfrac{a}{2b}$
f) $\quad\dfrac{2ac}{b-c}$
g) $\quad(2c)^2$
h) $\quad2c^2$

### Answers

a) $\quad12$
b) $\quad11$
c) $\quad-24$
d) $\quad-3$
e) $\quad-2$
f) $\quad-12$
g) $\quad36$
h) $\quad18$

## Hide

# Solving Equations [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/b_solving-equations.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* develop and solve linear equations, including those derived from substituting values into a formula, or those developed from a word description

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Equations

::: {.definition}

**Definition 1 - Equation**

A statement asserting that two *expressions* are equal.

:::

For example:

$$\underset{\overset{\uparrow}{\text{expression 1}}}{3n-3}\overset{\underset{\downarrow}{\text{is equal to}}}{=}\underset{\overset{\uparrow}{\text{expression 2}}}{n+1}$$

### Solving Equations

::: {.definition}

**Definition 2**

To **solve** an equation means to find the value of the unknown.

:::

In order to solve an equation we need to rearrange the equation to **isolate** the unknown. We do this by unpacking the equation using **inverse operations**. The inverse operations are performed on *both sides* of the equation to **maintain the balance**.

Once you have found a solution, you can check it is correct by **substituting** back into the original equation.

#### {.tabset .tabset-fade .tabset-pills}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

:::

###### Solution (a)

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

::: {.solution}

$\begin{aligned}2x-3&=5\\2x-3\color{red}{+3}&=5\color{red}{+3}\\2x&=8\\2x\color{red}{\div2}&=8\color{red}{\div2}\\x&=4\end{aligned}$

:::

###### Solution (b)

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

::: {.solution}

$\begin{aligned}8-4x&=-2\\8-4x\color{red}{+4x+2}&=-2\color{red}{+4x+2}\\10&=4x\\10\color{red}{\div4}&=4x\color{red}{\div4}\\\frac{10}{4}&=x\\x&=\frac{10}{4}\\x&=\frac{5}{2}=2.5\end{aligned}$

:::

::: {.note}

It is best practice to have the unknown on the left hand side (LHS) in your answer.

:::

###### Solution (c)

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

::: {.solution}

$\begin{aligned}\frac{x}{4}+7&=5\\\frac{x}{4}+7\color{red}{-7}&=5\color{red}{-7}\\\frac{x}{4}&=-2\\\frac{x}{4}\color{red}{\times4}&=-2\color{red}{\times4}\\x&=-8\end{aligned}$

:::

###### Solution (d)

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

::: {.solution}

$\begin{aligned}\frac{1}{3}(x+2)&=6\\\frac{1}{3}(x+2)\color{red}{\times3}&=6\color{red}{\times3}\\x+2&=18\\x+2\color{red}{-2}&=18\color{red}{-2}\\x&=16\end{aligned}$

:::

##### Hide

### Repeated Unknowns

If the unknown appears in the equation more than once, we follow these steps:
1. Expand brackets and collect like terms.
2. If the unknown appears on both sides of the equation, remove it from one side using inverse operations.
3. Isolate the unknown and solve the equation.

#### {.tabset .tabset-fade .tabset-pills}

##### Example 2 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve $\quad4(2x+5)-3(x-2)=16\quad$ for $x$.

:::

###### Solution

Solve $\quad4(2x+5)-3(x-2)=16\quad$ for $x$.

::: {.solution}

$\begin{aligned}4(2x+5)-3(x-2)&=16\\8x+20-3x+6&=16\\5x+26&=16\\5x+26\color{red}{-26}&=16\color{red}{-26}\\5x&=-10\\5x\color{red}{\div5}&=-10\color{red}{\div5}\\x&=-2\end{aligned}$

:::

##### Example 3 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve for $x$:

a) $\quad4x-3=3x+7$
b) $\quad5-3(-1+x)=x$

:::

###### Solution (a)

Solve for $x$:

a) $\quad4x-3=3x+7$
b) $\quad5-3(-1+x)=x$

::: {.solution}

$\begin{aligned}4x-3&=3x+7\\4x\color{red}{-3x}-3\color{red}{+3}&=3x\color{red}{-3x}+7\color{red}{+3}\\x&=10\end{aligned}$

:::

###### Solution (b)

Solve for $x$:

a) $\quad4x-3=3x+7$
b) $\quad5-3(-1+x)=x$

::: {.solution}

$\begin{aligned}5-3(-1+x)&=x\\5+3-3x&=x\\8-3x\color{red}{+3x}&=x\color{red}{+3x}\\8&=4x\\8\color{red}{\div4}&=4x\color{red}{\div4}\\2&=x\\x&=2\end{aligned}$

:::

##### Hide

### Fractions

For equations involving fractions, first write all fractions with a **common denominator**.

For example:

* In $\ \ \dfrac{2x}{3}=\dfrac{x}{4}\ \ $ the (lowest) common denominator is $\ \ 12$.
* In $\ \ \dfrac{3}{x}=\dfrac{5}{2x-1}\ \ $ the common denominator is $\ \ x(2x-1)$.

Once the fractions have common denominators you can **equate the numerators**

#### {.tabset .tabset-fade .tabset-pills}

##### Example 4 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve $\quad\dfrac{2-x}{3}=\dfrac{x}{5}\quad$ for $x$.

:::

###### Solution

Solve $\quad\dfrac{2-x}{3}=\dfrac{x}{5}\quad$ for $x$.

::: {.solution}

$\begin{aligned}\frac{2-x}{3}&=\frac{x}{5}\\\frac{2-x}{3}\color{red}{\times\frac{5}{5}}&=\frac{x}{5}\color{red}{\times\frac{3}{3}}\\\frac{5(2-x)}{15}&=\frac{3x}{15}\\5(2-x)&=3x&\text{(equating the numerators)}\\10-5x&=3x\\10-5x\color{red}{+5x}&=3x\color{red}{+5x}\\10&=8x\\10\color{red}{\div8}&=8x\color{red}{\div8}\\\frac{10}{8}&=x\\x&=\frac{10}{8}\\x&=\frac{5}{4}=1.25\end{aligned}$

:::

##### Example 5 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve $\quad\dfrac{7}{x+1}=\dfrac{3}{x}\quad$ for $x$.

:::

###### Solution

Solve $\quad\dfrac{7}{x+1}=\dfrac{3}{x}\quad$ for $x$.

::: {.solution}

$\begin{aligned}\frac{7}{x+1}&=\frac{3}{x}\\\frac{7}{x+1}\color{red}{\times\frac{x}{x}}&=\frac{3}{x}\color{red}{\times\frac{x+1}{x+1}}\\\frac{7x}{x(x+1)}&=\frac{3(x+1)}{x(x+1)}\\7x&=3(x+1)&\text{(equating the numerators)}\\7x&=3x+3\\7x\color{red}{-3x}&=3x\color{red}{-3x}+3\\4x&=3\\4x\color{red}{\div4}&=3\color{red}{\div4}\\x&=\frac{3}{4}=0.75\\\end{aligned}$

:::

##### Hide

## Hide

# Formulae [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/c_formulae.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* review evaluating the subject of a formula, given the value of other pronumerals in the formula

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Formulae

::: {.definition}

A **formula** is an equation which connects two or more variables. The plural of formula is **formulae** or **formulas**.

:::

Generally a formula is written with a single variable, the **subject** on the LHS and the rest on the right.

$$\underset{\overset{\uparrow}{\text{subject}}}{A}=\frac{1}{2}ab\cos{C}$$

### Formula substitution

If the formula contains two or more variables and we know the value of all but one of them, we can solve an equation to find the remaining variable.

1) State the variables in the formula and the known values.
2) Write down the formula and Substitute the known values into the formula.
3) Solve the one variable equation to find the unknown value.

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

The area of a trapezium is given by $\ \ A=\dfrac{h}{2}(a+b)\ \ $ where $a$ and $b$ are the lengths of the parallel sides and $h$ is the height.

Calculate the height of a 20 cm$^2$ trapezium with parallel lengths of 3 cm and 7 cm.

:::

###### Solution

The area of a trapezium is given by $\ \ A=\dfrac{h}{2}(a+b)\ \ $ where $a$ and $b$ are the lengths of the parallel sides and $h$ is the height.

Calculate the height of a 20 cm$^2$ trapezium with parallel lengths of 3 cm and 7 cm.

::: {.solution}

<div class="row">
  <div class="column">
  $\begin{aligned}A&=20\text{ cm}^2\\h&=\ ?\\a&=3\text{ cm}^2\\b&=7\text{ cm}^2\end{aligned}$
  </div>
  <div class="column">
  $\begin{aligned}A&=\frac{h}{2}(a+b)\\20&=\frac{h}{2}(3+7)\\20&=\frac{h}{2}(10)\\20&=5h\\20\color{red}{\div5}&=5h\color{red}{\div5}\\4&=h\\h&=4\text{ cm}\end{aligned}$
  </div>
</div>

The height of the trapezium is equal to 4 cm.

:::

##### Hide

## Hide
# Rearranging Formulae [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/d_rearranging-formulae.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* change the subject of a formula

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Formula Rearrangement

By performing operations to both sides of a formula, they can be **rearranged** to make **equivalent** formulae where other variables are the subjects.

<div class="row">
  <div class="column">
  $$\begin{aligned}V&=\pi r^2h\\V\color{red}{\div\pi r^2}&=\pi r^2h\color{red}{\div\pi r^2}\\\frac{V}{\pi r^2}&=h\\h&=\frac{V}{\pi r^2}\end{aligned}$$
  </div>
  <div class="column">
  $$\begin{aligned}V&=\pi r^2h\\V\color{red}{\div\pi h}&=\pi r^2h\color{red}{\div\pi h}\\\frac{V}{\pi h}&=r^2\\r^2&=\frac{V}{\pi h}\\\color{red}{\sqrt{\color{black}{r^2}}}&=\color{red}{\sqrt{\color{black}{\frac{V}{\pi h}}}}\\r&=\sqrt{\frac{V}{\pi h}}&(r>0)\end{aligned}$$
  </div>
</div>

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Make $a$ the subject of $\quad2a-7b=23$.

:::

###### Solution

Make $a$ the subject of $\quad2a-7b=23$.

::: {.solution}

$\begin{aligned}2a-7b&=23\\2a-7b\color{red}{+7b}&=23\color{red}{+7b}\\2a&=23-7b\\\color{red}{\frac{\color{black}{2a}}{2}}&=\color{red}{\frac{\color{black}{23-7b}}{2}}\\a&=\frac{23-7b}{2}\end{aligned}$

:::

##### Hide

### Rearranging and substituting

Previously, during formula substitution, the variables were replaced by numbers and then the equation was solved. However, often we need to substitute several values for the unknowns and solve the equation for each case. In this situation it is quicker to **rearrange** the formula **before substituting**.

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 2 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

The surface area of a sphere is given by $\ \ A=4\pi r^2\ \ $ where $r$ is the sphere's radius.

a) Rearrange this formula to make $r$ the subject.
b) Hence find the radius to 3 significant figures when the surface area is:
    i) 10 cm$^2$
    ii) 20 cm$^2$
    iii) 30 cm$^3$

:::

###### Solution (a)

The surface area of a sphere is given by $\ \ A=4\pi r^2\ \ $ where $r$ is the sphere's radius.

a) Rearrange this formula to make $r$ the subject.
b) Hence find the radius to 3 significant figures when the surface area is:
    i) 10 cm$^2$
    ii) 20 cm$^2$
    iii) 30 cm$^3$

::: {.solution}

$\begin{aligned}A&=4\pi r^2\\A\color{red}{\div4\pi}&=4\pi r^2\color{red}{\div4\pi}\\\frac{A}{4\pi}&=r^2\\\color{red}{\sqrt{\color{black}{\frac{A}{4\pi}}}}&=\color{red}{\sqrt{\color{black}{r^2}}}\\r&=\sqrt{\frac{A}{4\pi}}&(r>0)\end{aligned}$

:::

###### Solution (b)

The surface area of a sphere is given by $\ \ A=4\pi r^2\ \ $ where $r$ is the sphere's radius.

a) Rearrange this formula to make $r$ the subject.
b) Hence find the radius to 3 significant figures when the surface area is:
    i) 10 cm$^2$
    ii) 20 cm$^2$
    iii) 30 cm$^3$

::: {.solution}

$\begin{aligned}\text{i) }\ \ r&=\sqrt{\frac{A}{4\pi}}&\text{ii) }\ \ r&=\sqrt{\frac{A}{4\pi}}&\text{iii) }\ \ r&=\sqrt{\frac{A}{4\pi}}\\\\r&=\sqrt{\frac{\color{red}{10}}{4\pi}}&r&=\sqrt{\frac{\color{red}{20}}{4\pi}}&r&=\sqrt{\frac{\color{red}{30}}{4\pi}}\\\\r&=0.892\text{ cm}&r&=1.26\text{ cm}&r&=1.55\text{ cm}\end{aligned}$

:::

##### Hide

## Hide