### 2021-01-26 ms-a1_formulae-and-equations.rmd
Added MS-A1_e, MS-A1_f and MS-A1_g. Added calc font, added formula box

---
title: "MS-A1 Formulae and Equations"
output:
  html_document:
    toc: true
    toc_float: true
---

# Syllabus [![](open.svg){width=16}](https://syllabus.nesa.nsw.edu.au/assets/mathematics_standard/mathematics-standard-stage-6-syllabus-2017.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

Students:

* review substitution of numerical values into linear and non-linear algebraic expressions and equations
  + review evaluating the subject of a formula, given the value of other pronumerals in the formula
  + change the subject of a formula
  + solve problems involving formulae, including calculating distance, speed and time (with change of units of measurement as required) or calculating stopping distances of vehicles using a suitable formula
* develop and solve linear equations, including those derived from substituting values into a formula, or those developed from a word description
* calculate and interpret blood alcohol content (BAC) based on drink consumption and body weight
  + use formulae, both in word form and algebraic form, to calculate an estimate for blood alcohol content $(BAC)$, including $BAC_{Male}=\frac{10N-7.5H}{6.8M}$ and $BAC_{Female}=\frac{10N-7.5H}{5.5M}$ where $N$ is the number of standard drinks consumed, $H$ is the number of hours of drinking, and $M$ is the person's weight in kilograms.
  + determine the number of hours required for a person to stop consuming alcohol in order to reach zero BAC, eg using the formula $time=\frac{BAC}{0.015}$
  + describe the limitations of methods estimating BAC
* calculate required medication dosages for children and adults from packets, given age or weight, using Fried’s, Young’s or Clark’s formula as appropriate
  + Fried's formula: Dosage for children 1-2 years $=\frac{\text{age (in months)}\ \times\ \text{adult dosage}}{150}$
  + Young's formula: Dosage for children 1-12 years $=\frac{\text{age of child (in years)}\ \times\ \text{adult dosage}}{\text{age of child (in years)}\ +\ 150}$
  + Clark's formula: Dosage $=\frac{\text{weight in kg}\ \times\ \text{adult dosage}}{70}$

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

## Hide

# Substitution [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/a_substitution.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* review substitution of numerical values into linear and non-linear algebraic expressions and equaitons

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Why algebra?

**Algebra** is used Mathematicians to communicate mathematical ideas in a convenient way. **Variables** (letters and symbols) are used to represent unknown quantities whose value can change depending on the situation.

### Expressions

We can think of expressions as number crunching machines. Numbers are put in and then a related number is produced.

For example the expression $a^2+1$ starts with an input $a$, squares it and adds 1.

If the number -5 is fed into the machine, the machine **substitutes** the number 5 in place of a and then **evaluates** the result:

$$\begin{aligned}&a^2+1\\=\ &(\color{red}{-5})^2+1\\=\ &25+1\\ =\ &26\end{aligned}$$

::: {.important}

**Important Note**

To avoid making mistakes it is good practice to **always** substitute into brackets.

:::

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Evaluate the following, given $x=4$

a) $\quad2x-3$
b) $\quad\sqrt{6x+12}$

:::

###### Solution (a)

Evaluate the following, given $x=4$

a) $\quad2x-3$
b) $\quad\sqrt{6x+12}$

::: {.solution}

$\begin{aligned}&2x-3\\=\ &2(\color{red}{4})-3\\=\ &8-3\\=\ &5\end{aligned}$

:::

###### Solution (b)

Evaluate the following, given $x=4$

a) $\quad2x-3$
b) $\quad\sqrt{6x+12}$

::: {.solution}

$\begin{aligned}&\sqrt{6x+12}\\=\ &\sqrt{6(\color{red}{4})+12}\\=\ &\sqrt{24+12}\\=\ &\sqrt{36}\\=\ &6\end{aligned}$

:::

##### Example 2 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Evaluate $\quad a^2-5b+c\quad$ given $a=3$, $b=-3$, $c=7$

:::

###### Solution

Evaluate $\quad a^2-5b+c\quad$ given $a=3$, $b=-3$, $c=7$

::: {.solution}

$\begin{aligned}&a^2-5b+c\\=\ &(\color{red}{3})^2-5(\color{red}{-3})+(\color{red}{7})\\=\ &9+15+7\\=\ &31\end{aligned}$

:::

##### Hide

## Practice {.tabset .tabset-fade .tabset-pills}

### Questions

If $a=4$, $b=-1$, and $c=-3$, evaluate:

a) $\quad3a$
b) $\quad2a-3b$
c) $\quad2ac$
d) $\quad ac-3c$
e) $\quad\dfrac{a}{2b}$
f) $\quad\dfrac{2ac}{b-c}$
g) $\quad(2c)^2$
h) $\quad2c^2$

### Answers

a) $\quad12$
b) $\quad11$
c) $\quad-24$
d) $\quad-3$
e) $\quad-2$
f) $\quad-12$
g) $\quad36$
h) $\quad18$

## Hide

# Solving Equations [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/b_solving-equations.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* develop and solve linear equations, including those derived from substituting values into a formula, or those developed from a word description

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Equations

::: {.definition}

**Definition 1 - Equation**

A statement asserting that two *expressions* are equal.

:::

For example:

$$\underset{\overset{\uparrow}{\text{expression 1}}}{3n-3}\overset{\underset{\downarrow}{\text{is equal to}}}{=}\underset{\overset{\uparrow}{\text{expression 2}}}{n+1}$$

### Solving Equations

::: {.definition}

**Definition 2**

To **solve** an equation means to find the value of the unknown.

:::

In order to solve an equation we need to rearrange the equation to **isolate** the unknown. We do this by unpacking the equation using **inverse operations**. The inverse operations are performed on *both sides* of the equation to **maintain the balance**.

Once you have found a solution, you can check it is correct by **substituting** back into the original equation.

#### {.tabset .tabset-fade .tabset-pills}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

:::

###### Solution (a)

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

::: {.solution}

$\begin{aligned}2x-3&=5\\2x-3\color{red}{+3}&=5\color{red}{+3}\\2x&=8\\2x\color{red}{\div2}&=8\color{red}{\div2}\\x&=4\end{aligned}$

:::

###### Solution (b)

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

::: {.solution}

$\begin{aligned}8-4x&=-2\\8-4x\color{red}{+4x+2}&=-2\color{red}{+4x+2}\\10&=4x\\10\color{red}{\div4}&=4x\color{red}{\div4}\\\frac{10}{4}&=x\\x&=\frac{10}{4}\\x&=\frac{5}{2}=2.5\end{aligned}$

:::

::: {.note}

It is best practice to have the unknown on the left hand side (LHS) in your answer.

:::

###### Solution (c)

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

::: {.solution}

$\begin{aligned}\frac{x}{4}+7&=5\\\frac{x}{4}+7\color{red}{-7}&=5\color{red}{-7}\\\frac{x}{4}&=-2\\\frac{x}{4}\color{red}{\times4}&=-2\color{red}{\times4}\\x&=-8\end{aligned}$

:::

###### Solution (d)

Solve for $x$:

a) $\quad2x-3=5$
b) $\quad8=4x=-2$
c) $\quad\dfrac{x}{4}+7=5$
d) $\quad\dfrac{1}{3}(x+2)=6$

::: {.solution}

$\begin{aligned}\frac{1}{3}(x+2)&=6\\\frac{1}{3}(x+2)\color{red}{\times3}&=6\color{red}{\times3}\\x+2&=18\\x+2\color{red}{-2}&=18\color{red}{-2}\\x&=16\end{aligned}$

:::

##### Hide

### Repeated Unknowns

If the unknown appears in the equation more than once, we follow these steps:
1. Expand brackets and collect like terms.
2. If the unknown appears on both sides of the equation, remove it from one side using inverse operations.
3. Isolate the unknown and solve the equation.

#### {.tabset .tabset-fade .tabset-pills}

##### Example 2 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve $\quad4(2x+5)-3(x-2)=16\quad$ for $x$.

:::

###### Solution

Solve $\quad4(2x+5)-3(x-2)=16\quad$ for $x$.

::: {.solution}

$\begin{aligned}4(2x+5)-3(x-2)&=16\\8x+20-3x+6&=16\\5x+26&=16\\5x+26\color{red}{-26}&=16\color{red}{-26}\\5x&=-10\\5x\color{red}{\div5}&=-10\color{red}{\div5}\\x&=-2\end{aligned}$

:::

##### Example 3 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve for $x$:

a) $\quad4x-3=3x+7$
b) $\quad5-3(-1+x)=x$

:::

###### Solution (a)

Solve for $x$:

a) $\quad4x-3=3x+7$
b) $\quad5-3(-1+x)=x$

::: {.solution}

$\begin{aligned}4x-3&=3x+7\\4x\color{red}{-3x}-3\color{red}{+3}&=3x\color{red}{-3x}+7\color{red}{+3}\\x&=10\end{aligned}$

:::

###### Solution (b)

Solve for $x$:

a) $\quad4x-3=3x+7$
b) $\quad5-3(-1+x)=x$

::: {.solution}

$\begin{aligned}5-3(-1+x)&=x\\5+3-3x&=x\\8-3x\color{red}{+3x}&=x\color{red}{+3x}\\8&=4x\\8\color{red}{\div4}&=4x\color{red}{\div4}\\2&=x\\x&=2\end{aligned}$

:::

##### Hide

### Fractions

For equations involving fractions, first write all fractions with a **common denominator**.

For example:

* In $\ \ \dfrac{2x}{3}=\dfrac{x}{4}\ \ $ the (lowest) common denominator is $\ \ 12$.
* In $\ \ \dfrac{3}{x}=\dfrac{5}{2x-1}\ \ $ the common denominator is $\ \ x(2x-1)$.

Once the fractions have common denominators you can **equate the numerators**

#### {.tabset .tabset-fade .tabset-pills}

##### Example 4 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve $\quad\dfrac{2-x}{3}=\dfrac{x}{5}\quad$ for $x$.

:::

###### Solution

Solve $\quad\dfrac{2-x}{3}=\dfrac{x}{5}\quad$ for $x$.

::: {.solution}

$\begin{aligned}\frac{2-x}{3}&=\frac{x}{5}\\\frac{2-x}{3}\color{red}{\times\frac{5}{5}}&=\frac{x}{5}\color{red}{\times\frac{3}{3}}\\\frac{5(2-x)}{15}&=\frac{3x}{15}\\5(2-x)&=3x&\text{(equating the numerators)}\\10-5x&=3x\\10-5x\color{red}{+5x}&=3x\color{red}{+5x}\\10&=8x\\10\color{red}{\div8}&=8x\color{red}{\div8}\\\frac{10}{8}&=x\\x&=\frac{10}{8}\\x&=\frac{5}{4}=1.25\end{aligned}$

:::

##### Example 5 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Solve $\quad\dfrac{7}{x+1}=\dfrac{3}{x}\quad$ for $x$.

:::

###### Solution

Solve $\quad\dfrac{7}{x+1}=\dfrac{3}{x}\quad$ for $x$.

::: {.solution}

$\begin{aligned}\frac{7}{x+1}&=\frac{3}{x}\\\frac{7}{x+1}\color{red}{\times\frac{x}{x}}&=\frac{3}{x}\color{red}{\times\frac{x+1}{x+1}}\\\frac{7x}{x(x+1)}&=\frac{3(x+1)}{x(x+1)}\\7x&=3(x+1)&\text{(equating the numerators)}\\7x&=3x+3\\7x\color{red}{-3x}&=3x\color{red}{-3x}+3\\4x&=3\\4x\color{red}{\div4}&=3\color{red}{\div4}\\x&=\frac{3}{4}=0.75\\\end{aligned}$

:::

##### Hide

## Hide

# Formulae [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/c_formulae.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* review evaluating the subject of a formula, given the value of other pronumerals in the formula

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Formulae

::: {.definition}

A **formula** is an equation which connects two or more variables. The plural of formula is **formulae** or **formulas**.

:::

Generally a formula is written with a single variable, the **subject** on the LHS and the rest on the right.

$$\underset{\overset{\uparrow}{\text{subject}}}{A}=\frac{1}{2}ab\cos{C}$$

### Formula substitution

If the formula contains two or more variables and we know the value of all but one of them, we can solve an equation to find the remaining variable.

1) State the variables in the formula and the known values.
2) Write down the formula and Substitute the known values into the formula.
3) Solve the one variable equation to find the unknown value.

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

The area of a trapezium is given by $\ \ A=\dfrac{h}{2}(a+b)\ \ $ where $a$ and $b$ are the lengths of the parallel sides and $h$ is the height.

Calculate the height of a 20 cm$^2$ trapezium with parallel lengths of 3 cm and 7 cm.

:::

###### Solution

The area of a trapezium is given by $\ \ A=\dfrac{h}{2}(a+b)\ \ $ where $a$ and $b$ are the lengths of the parallel sides and $h$ is the height.

Calculate the height of a 20 cm$^2$ trapezium with parallel lengths of 3 cm and 7 cm.

::: {.solution}

<div class="row">
  <div class="column">
  $\begin{aligned}A&=20\text{ cm}^2\\h&=\ ?\\a&=3\text{ cm}^2\\b&=7\text{ cm}^2\end{aligned}$
  </div>
  <div class="column">
  $\begin{aligned}A&=\frac{h}{2}(a+b)\\20&=\frac{h}{2}(3+7)\\20&=\frac{h}{2}(10)\\20&=5h\\20\color{red}{\div5}&=5h\color{red}{\div5}\\4&=h\\h&=4\text{ cm}\end{aligned}$
  </div>
</div>

The height of the trapezium is equal to 4 cm.

:::

##### Hide

## Hide
# Rearranging Formulae [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/d_rearranging-formulae.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* change the subject of a formula

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Formula Rearrangement

By performing operations to both sides of a formula, they can be **rearranged** to make **equivalent** formulae where other variables are the subjects.

<div class="row">
  <div class="column">
  $$\begin{aligned}V&=\pi r^2h\\V\color{red}{\div\pi r^2}&=\pi r^2h\color{red}{\div\pi r^2}\\\frac{V}{\pi r^2}&=h\\h&=\frac{V}{\pi r^2}\end{aligned}$$
  </div>
  <div class="column">
  $$\begin{aligned}V&=\pi r^2h\\V\color{red}{\div\pi h}&=\pi r^2h\color{red}{\div\pi h}\\\frac{V}{\pi h}&=r^2\\r^2&=\frac{V}{\pi h}\\\color{red}{\sqrt{\color{black}{r^2}}}&=\color{red}{\sqrt{\color{black}{\frac{V}{\pi h}}}}\\r&=\sqrt{\frac{V}{\pi h}}&(r>0)\end{aligned}$$
  </div>
</div>

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Make $a$ the subject of $\quad2a-7b=23$.

:::

###### Solution

Make $a$ the subject of $\quad2a-7b=23$.

::: {.solution}

$\begin{aligned}2a-7b&=23\\2a-7b\color{red}{+7b}&=23\color{red}{+7b}\\2a&=23-7b\\\color{red}{\frac{\color{black}{2a}}{2}}&=\color{red}{\frac{\color{black}{23-7b}}{2}}\\a&=\frac{23-7b}{2}\end{aligned}$

:::

##### Hide

### Rearranging and substituting

Previously, during formula substitution, the variables were replaced by numbers and then the equation was solved. However, often we need to substitute several values for the unknowns and solve the equation for each case. In this situation it is quicker to **rearrange** the formula **before substituting**.

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 2 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

The surface area of a sphere is given by $\ \ A=4\pi r^2\ \ $ where $r$ is the sphere's radius.

a) Rearrange this formula to make $r$ the subject.
b) Hence find the radius to 3 significant figures when the surface area is:
    i) 10 cm$^2$
    ii) 20 cm$^2$
    iii) 30 cm$^3$

:::

###### Solution (a)

The surface area of a sphere is given by $\ \ A=4\pi r^2\ \ $ where $r$ is the sphere's radius.

a) Rearrange this formula to make $r$ the subject.
b) Hence find the radius to 3 significant figures when the surface area is:
    i) 10 cm$^2$
    ii) 20 cm$^2$
    iii) 30 cm$^3$

::: {.solution}

$\begin{aligned}A&=4\pi r^2\\A\color{red}{\div4\pi}&=4\pi r^2\color{red}{\div4\pi}\\\frac{A}{4\pi}&=r^2\\\color{red}{\sqrt{\color{black}{\frac{A}{4\pi}}}}&=\color{red}{\sqrt{\color{black}{r^2}}}\\r&=\sqrt{\frac{A}{4\pi}}&(r>0)\end{aligned}$

:::

###### Solution (b)

The surface area of a sphere is given by $\ \ A=4\pi r^2\ \ $ where $r$ is the sphere's radius.

a) Rearrange this formula to make $r$ the subject.
b) Hence find the radius to 3 significant figures when the surface area is:
    i) 10 cm$^2$
    ii) 20 cm$^2$
    iii) 30 cm$^3$

::: {.solution}

$\begin{aligned}\text{i) }\ \ r&=\sqrt{\frac{A}{4\pi}}&\text{ii) }\ \ r&=\sqrt{\frac{A}{4\pi}}&\text{iii) }\ \ r&=\sqrt{\frac{A}{4\pi}}\\\\r&=\sqrt{\frac{\color{red}{10}}{4\pi}}&r&=\sqrt{\frac{\color{red}{20}}{4\pi}}&r&=\sqrt{\frac{\color{red}{30}}{4\pi}}\\\\r&=0.892\text{ cm}&r&=1.26\text{ cm}&r&=1.55\text{ cm}\end{aligned}$

:::

##### Hide

## Hide

# Speed, Distance and Time [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/e_speed-distance-time.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* review substitution of numerical values into linear and non-linear algebraic expressions and equations
  + solve problems involving formulae, including calculating distance, speed and time (with change of units of measurement as required) or calculating stopping distances of vehicles using a suitable formula

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Speed

::: {.definition}

**Definition**

**Speed** is a comparison between *distance travelled* and *time taken*

:::

In most real world situations, an objects speed is not constant. Therefore we often use the **average speed**.

::: {.formula}

$$s=\frac{d}{t}$$

Where:

* $s$ - average speed
* $d$ - distance travelled
* $t$ - time taken

::: {.important}

The formula is **not** on the reference sheet.

:::

:::

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Find the average speed of a car which travels 110 km in 2 hours.

:::

###### Solution

Find the average speed of a car which travels 110 km in 2 hours

::: {.solution}

$\begin{aligned}s&=\frac{d}{t}\\&=\frac{110\text{ km}}{2\text{ h}}\\&= 55\text{ km/h}\end{aligned}$

:::

##### Hide

### Rearranging

The formula for speed can be rewritten to make *distance* or *time* the subject.

$$s=\frac{d}{t}$$

$$d=st$$

$$t=\frac{d}{s}$$

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 2 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Jonah rides his motorcycle on a highway at an average speed of 90 km/h.

a) How far can Jonah travel in $1\frac{1}{2}$ hours?
b) How long will he take to travel 210 km? Answer in hours and minutes.

:::

###### Solution (a)

Jonah rides his motorcycle on a highway at an average speed of 90 km/h.

a) How far can Jonah travel in $1\frac{1}{2}$ hours?
b) How long will he take to travel 210 km? Answer in hours and minutes.

::: {.solution}

<div class="row">
  <div class="column">
  $\begin{aligned}d&=st\\d&=(90)\underset{\color{red}{(*)}}{\left(\color{red}{1\frac{1}{2}}\right)}\\d&=135\text{ km}\end{aligned}$
  </div>
  <div class="column">
  $(*)=$ <span style="font-family: casio">qa1$1R2</span>
  </div>
</div>

:::

###### Solution (b)

Jonah rides his motorcycle on a highway at an average speed of 90 km/h.

a) How far can Jonah travel in $1\frac{1}{2}$ hours?
b) How long will he take to travel 210 km? Answer in hours and minutes.

::: {.solution}

<div class="row">
  <div class="column">
  $\begin{aligned}t&=\frac{d}{s}\\t&=\frac{210}{90}\\t&=2.33...\text{ h}\\&\quad\color{red}{(*)}\\t&=2^\circ 20' 0''\end{aligned}$

  2 hours and 20 minutes
  </div>
  <div class="column">
  $(*)=$ <span style="font-family: casio">x</span>
  </div>
</div>

:::

##### Hide

### Stopping distance

::: {.definition}

**Definition - Stopping distance**

The distance a vehicle will travel in order to come to a complete stop.

:::

<center>stopping distance = reaction distance + braking distance</center>

::: {.important}

**Important Note**

You **are not** expected to memorise any formulae.

You **are** expected to be able to substitute into a given formula.

:::

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 3 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Claire is driving on a motorway at a speed of 110 kilometres per hour and has to brake suddenly. She has a reaction time of 2 seconds and a braking distance of 59.2 metres.

What is Claire's stopping distance?

:::

###### Solution

Claire is driving on a motorway at a speed of 110 kilometres per hour and has to brake suddenly. She has a reaction time of 2 seconds and a braking distance of 59.2 metres.

What is Claire's stopping distance?

::: {.solution}

$\begin{aligned}110\text{ km/h}&=110000\text{ m/hr}\\&=\frac{110000}{60\times60}\text{ m/s}\\&=30.555...\text{ m/s}\end{aligned}$

Let $d$ be Claire's stopping distance.

$\begin{aligned}d&=\underbrace{\text{reaction time distance}}_{d=st}+\text{braking distance}\\&=st+\text{braking distance}\\&=(30.555...)(2)+(59.2)\\&=120.311...\\d&=120.3 \text{ m (1 d.p.)}\end{aligned}$

Claire's stopping distance is 120.3 metres.

:::

##### Example 4 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Max was driving 60 km/h and has a reaction time of 0.8 s. Calculate the stopping distance correct to the nearest metre given the formula $\quad d=\dfrac{5vt}{18}+\dfrac{v^2}{170}\quad$ where $d$ is the stopping distance (m), $v$ is the speed (m/s) and $t$ is reaction time (s).

:::

###### Solution

Max was driving 60 km/h and has a reaction time of 0.8 s. Calculate the stopping distance correct to the nearest metre given the formula $\quad d=\dfrac{5vt}{18}+\dfrac{v^2}{170}\quad$ where $d$ is the stopping distance (m), $v$ is the speed (m/s) and $t$ is reaction time (s).

::: {.solution}

\begin{aligned}d&=\frac{5vt}{18}+\frac{v^2}{170}\\d&=\frac{5(60)(0.8)}{18}+\frac{60^2}{170}\\d&=34.509...\\d&\approx35\text{ m}\end{aligned}

Max's stopping distance is 35 metres.

:::

##### Hide

## Hide

# Blood Alcohol Content [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/f_blood-alcohol-content.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* calculate and interpret blood alcohol content (BAC) based on drink consumption and body weight
  + use formulae, both in word form and algebraic form, to calculate an estimate for blood alcohol content $(BAC)$, including $BAC_{Male}=\frac{10N-7.5H}{6.8M}$ and $BAC_{Female}=\frac{10N-7.5H}{5.5M}$ where $N$ is the number of standard drinks consumed, $H$ is the number of hours of drinking, and $M$ is the person's weight in kilograms.
  + determine the number of hours required for a person to stop consuming alcohol in order to reach zero BAC, eg using the formula $time=\frac{BAC}{0.015}$
  + describe the limitations of methods estimating BAC

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Blood Alcohol Content

::: {.definition}

**Definition**

Blood alcohol content (BAC) is a measure of the amount of alcohol in your blood.

:::

BAC tells you the number of grams of alcohol per 100 mL of blood

For example, a BAC of 0.10 means there is 0.10 g of alcohol for every 100 mL of blood, or 1 g for every 1 L of blood.

### Estimating BAC

BAC can be estimated using the following formulae:

::: {.formula}

$$BAC_{\text{Male}}=\frac{10N-7.5H}{6.8M}\quad\text{or}\quad BAC_{\text{Female}}=\frac{10N-7.5H}{5.5M}$$

Where:
* $BAC$ - Blood alcohol content
* $N$ - Number of standard drinks consumed
* $H$ - Hours drinking
* $M$ - Mass in kilograms

::: {.important}

These formulae are **not** on the reference sheet.

You are **not** expected to remember these formulae.

:::

:::

### Limitations

These formulae are only good for calculating *estimates* of a person BAC.

The formulae for estimating BAC only takes into account the number of standard drinks consumed, sex, the amount of time, and mass.

There are other factors that influence BAC including:

* Fitness
* Health
* Liver function
* Food in the stomach
* Medications

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Osman is 87 kg and has consumed 3 standard drinks in the past hour. Estimate Osman's BAC to 3 decimak places using the following formula where $N$ represents the number of standard drinks, $H$ is the number of hours drinking and $M$ is his mass in kilograms.

$$BAC_{\text{Male}}=\frac{10N-7.5H}{6.8M}$$

:::

###### Solution

Osman is 87 kg and has consumed 3 standard drinks in the past hour. Estimate Osman's BAC to 3 decimak places using the following formula where $N$ represents the number of standard drinks, $H$ is the number of hours drinking and $M$ is his mass in kilograms.

$$BAC_{\text{Male}}=\frac{10N-7.5H}{6.8M}$$

::: {.solution}

$\begin{aligned}BAC_{\text{Male}}&=\frac{10N-7.5H}{6.8M}\\&=\frac{10(3)-7.5(1)}{6.8(87)}\\&=0.038032...\\&\approx0.038\end{aligned}$

:::

##### Hide

### Reaching zero BAC

The number of hours required for a person to reach zero BAC after they stop consuming alcohol is given by the following formula:

::: {.formula}

$$\text{Time (h)}=\frac{BAC}{0.015}$$

::: {.important}

This formula is **not** on the reference sheet.

You are **not** expected to remember this formula.
:::

:::

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 2 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Fiona has a BAC of 0.027. Use the following formula to estimate the time she will need for her BAC to reach zero. Answer to the nearest minute.

$$\text{Time}=\frac{BAC}{0.015}$$

:::

###### Solution

Fiona has a BAC of 0.027. Use the following formula to estimate the time she will need for her BAC to reach zero. Answer to the nearest minute.

$$\text{Time}=\frac{BAC}{0.015}$$

::: {.solution}

<div class="row">
  <div class="column">
  $\begin{aligned}\text{Time}&=\frac{BAC}{0.015}\\&=\frac{(0.027)}{0.015}\\&=1.8\\&\quad(*)\\&=1^\circ48'0''\end{aligned}$

  Fiona needs to wait 1 hour and 48 minutes.
  </div>
  <div class="column">
  $(*)=$ <span style="font-family: casio">x</span>
  </div>
</div>

:::

##### Hide

## Hide

# Medication [![](pres.svg){width=16}](https://claytonpollard.github.io/stage6_standard_presentations/MS-A1/g_medication.pdf){target="blank"} {.tabset .tabset-fade .tabset-pills}

## Show

::: {.syllabus}

**Syllabus**

* calculate required medication dosages for children and adults from packets, given age or weight, using Fried’s, Young’s or Clark’s formula as appropriate
  + Fried's formula: Dosage for children 1-2 years $=\frac{\text{age (in months)}\ \times\ \text{adult dosage}}{150}$
  + Young's formula: Dosage for children 1-12 years $=\frac{\text{age of child (in years)}\ \times\ \text{adult dosage}}{\text{age of child (in years)}\ +\ 150}$
  + Clark's formula: Dosage $=\frac{\text{weight in kg}\ \times\ \text{adult dosage}}{70}$

<div align="right"><i><small>As of 2021/01/20</small></i></div>
:::

### Converting concentrations

A concentration is a rate comparing a mass (g, mg, etc.) with a volume (L, mL, etc.). We need to consider both when converting a concentration from one unit to another.

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 1 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

The concentration of a mild analgesic is given as 80 mg per 50 mL. What is in g/mL?

:::

###### Solution

The concentration of a mild analgesic is given as 80 mg per 50 mL. What is in g/mL?

::: {.solution}

$\begin{aligned}80\text{ mg}/50\text{ mL}&=\frac{80\text{ mg}}{50\text{ mL}}\\&=1.6\text{ mg/mL}\\&=1.6\div1000\text{ g/mL}\\&=0.0016\text{ g/mL}\end{aligned}$

:::

##### Hide

### Calculating dosages

Often when calculating dosages you will be prescribed an amount in milligrams (mg) and need to calculate how much of a liquid that need to be taken. In these cases you can use the following:

::: {.formula}

$$\text{volume required}=\frac{\text{strength required}}{\text{strength of stock}}\times\text{volume of stock}$$

Where your *stock* is the medication you're given.

::: {.important}

This formula is **not** on the reference sheet.

:::

:::

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 2 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

A patient is prescribed 1000 mg of a mild painkiller. The medication available contains 100 mg in 5 mL. How much medication should be given to the patient?

:::

###### Solution

A patient is prescribed 1000 mg of a mild painkiller. The medication available contains 100 mg in 5 mL. How much medication should be given to the patient?

::: {.solution}

$\begin{aligned}\text{volume required}&=\frac{\text{strength required}}{\text{strength of stock}}\times\text{volume of stock}\\&=\frac{(1000\text{ mg})}{(100\text{ mg})}\times(5\text{ mL})\\&=50\text{ mL}\end{aligned}$

:::

##### Hide

### Children and infants

There are three formulae used for children and infants:

::: {.formula}

* Fried's formula: Dosage for children 1-2 years $=\dfrac{\text{age (months)}\ \times\ \text{adult dosage}}{150}$
* Young's formula: Dosage for children 1-12 years $=\dfrac{\text{age of child (years)}\ \times\ \text{adult dosage}}{\text{age of child (years)}\ +\ 150}$
* Clark's formula: Dosage $=\dfrac{\text{weight (kg)}\ \times\ \text{adult dosage}}{70}$

::: {.important}

These formulae are **not** on the reference sheet.

You are **not** expected to remember these formulae.

:::

:::

#### {.tabset .tabset-fade .tabset-pills .unlisted .unnumbered}

##### Example 3 {.tabset .tabset-fade .tabset-pills}

###### Question

::: {.example}

Jessica is 6 months old. Use Fried's formula to find the required infant dose if the adult dost is 20 mL.

$$\text{Dosage}=\dfrac{\text{age (months)}\ \times\ \text{adult dosage}}{150}$$

:::

###### Solution

Jessica is 6 months old. Use Fried's formula to find the required infant dose if the adult dost is 20 mL.

$$\text{Dosage}=\dfrac{\text{age (months)}\ \times\ \text{adult dosage}}{150}$$

::: {.solution}

$\begin{aligned}\text{Dosage}&=\frac{\text{age (months)}\ \times\ \text{adult dosage}}{150}\\&=\frac{(6)\times(20)}{(150)}\\&=\frac{120}{150}\\&=0.8\text{ mL}\end{aligned}$

:::

##### Hide

## Hide